\section{Parameters and Primitives}

We define SHRINKS-SHA2-128s as a SHRINKS instance preferable for Bitcoin. It operates with the 
following set of parameters:

\begin{center}
\begin{tabular}{ c c c }
 Parameter & Value & Description \\ 
 \hline
 \texttt{H} & \texttt{SHA256} & The underlying hash function \\ 
 \texttt{n} & $128$ & Security parameter (bits) \\  
 \texttt{w} & $256$ & Winternitz parameter (8 bits per chain) \\
 \texttt{l}  & $16$ & Message digest length, the number of WOTS+C chains \\
 \texttt{Swn} & $2040$ & Target sum for WOTS+C \\
 \texttt{hsf} & $\color{red}x\color{black}$ & Maximum stateful tree height \\
 \texttt{hsl} & $\color{red}x\color{black}$ & Maximum stateless hypertree height \\
 \texttt{d} & $5$ & Stateless hypertree layers \\
 \texttt{a} & $14$ & FORS/PORS tree height \\
 \texttt{k} & $11$ & Number of FORS/PORS trees 
\end{tabular}
\end{center}

The foundation of SHRINKS is the hash function SHA-256, operated in specific modes to construct 
WOTS+C (Compressed Winternitz One-Time Signatures) and Unbalanced Merkle Trees.

\subsection{Hash Function Instantiation}

SHRINKS uses SHA-256 truncated to $n$ bits for all internal hash operations. 

\subsection{Tweakable Hash Function}

To prevent multi-target attacks and ensure domain separation between the various components 
(WOTS+, trees, stateless fallback, etc), we employ the Tweakable Hash Function construction defined in 
FIPS 205:

\begin{center}
\texttt{Tw (P,T,m) = H (P || T || m)[0:n]}
\end{center}
where: 
\begin{itemize}
    \item[] \texttt{P} is a public parameters
    \item[] \texttt{T} is a tweak  (context address)
    \item[] \texttt{m} is a message
\end{itemize}

This usage of tweaked hashing is critical. It ensures that a hash computed at one position in the 
Merkle tree cannot be substituted for a hash at another position, effectively mitigating generic 
tree-grafting attacks.

\subsubsection{Tweak Structure}

Tweaks are structured as follows to ensure uniqueness across all scheme components:

\begin{center}
\texttt{T = tweak\_type || tree\_address || chain\_index || position}
\end{center}
where: 
\begin{itemize}
    \item[] \texttt{tweak\_type} is type of the layer defined in Appendix A.
    \item[] \texttt{tree\_address} is a position in hypertree
    \item[] \texttt{chain\_index} is a WOTS chain index
    \item[] \texttt{position} is a position within a chain
\end{itemize}

\subsection{Pseudorandom Function}

\texttt{PRF(SK.seed) = H (0x06 || SK.seed)[0:n]}

\subsection{Message Hash Function}

\texttt{H\_msg(R, PK.seed, PK.root, m) = H(0x05 || R || PK.seed || PK.root || m)}

\subsection{WOTS+C}
The standard Winternitz One-Time Signature (WOTS+) is the building block of most modern hash-based 
schemes. It works by encoding the message into a series of small integers and using them to iterate 
hash chains. A significant portion of a WOTS+ signature size (approx 20\%) is dedicated to a 
"Checksum," which is necessary to prevent a forgery attack where an adversary extends the hash 
chains forward.

WOTS+C (Compressed WOTS+) eliminates the transmission of this checksum entirely, achieving a smaller 
signature size. In WOTS+, the checksum $C$ is calculated from the message digest $D$.

$$C = \sum_{i=1}^{\ell} (w - 1 - d_i)$$

This checksum is then signed using additional chains. In WOTS+C, instead of calculating a checksum 
for a fixed message digest, the signer modifies the message digest until it possesses a specific, 
pre-determined checksum property. This is done by appending a counter \texttt{count} and a random 
salt $R$ to the message before hashing. The signer increments cnt until the resulting hash digest 
$D$ satisfies a global target condition.

\subsubsection{The Target Condition}
For SHRINKS-SHA2-128s, we use the Winternitz parameter $w=256$. This means each chain processes 8 
bits (1 byte) of the digest. The message digest length is 128 bits (16 bytes), so $\ell = 16$. The 
values $d_i$ are treated as integers in $w$.

We define the expected value of the checksum. If the hash output is uniform, the expected value of a 
single digit is $(w-1)/2 = 127.5$. For 16 digits, the expected sum is $16 \times 127.5 = 2040$.

The condition for a valid WOTS+C message digest $D$ is:
$$\sum_{i=0}^{15} (255 - d_i) = 2040$$

By enforcing this condition during signing (including grinding), the verifier can simply assume 
the checksum is 2040 without receiving it. If the reconstructed digest does not sum to 2040, the 
signature is invalid. 

\subsection{UXMSS Tree}
Standard XMSS uses a balanced binary tree, which minimizes the maximum path length but imposes a 
logarithmic cost on every signature. For $N$ signatures, every signature has size proportional to 
$\log_2 N$. SHRINKS utilizes an Unbalanced Merkle Tree (specifically a "Right-Skewed" tree) to 
optimize for the typical Bitcoin use case: a wallet that performs very few transactions in its 
lifecycle (we are targeting $N < 10$).

In this topology the $k$-th signature requires a path of length $k+1$.This linear growth is superior 
to logarithmic growth for small $k$. For $k=0$, the path is just 16 bytes. Even for $k=10$, the path 
is $11 \times 16 = 176$ bytes, which is extremely small compared to the fixed ~2.5 KB of a balanced 
tree or the ~8 KB of SLH-DSA.